\subsection{fpminimax}
\label{labfpminimax}
\noindent Name: \textbf{fpminimax}\\
computes a good polynomial approximation with fixed-point or floating-point coefficients\\
\noindent Usage: 
\begin{center}
\textbf{fpminimax}(\emph{f}, \emph{n}, \emph{formats}, \emph{range}, \emph{indic1}, \emph{indic2}, \emph{indic3}, \emph{P}) : (\textsf{function}, \textsf{integer}, \textsf{list}, \textsf{range}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{function}) $\rightarrow$ \textsf{function}\\
\textbf{fpminimax}(\emph{f}, \emph{monomials}, \emph{formats}, \emph{range}, \emph{indic1}, \emph{indic2}, \emph{indic3}, \emph{P}) : (\textsf{function}, \textsf{list}, \textsf{list}, \textsf{range},  \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{function}) $\rightarrow$ \textsf{function}\\
\textbf{fpminimax}(\emph{f}, \emph{n}, \emph{formats}, \emph{L}, \emph{indic1}, \emph{indic2}, \emph{indic3}, \emph{P}) : (\textsf{function}, \textsf{integer}, \textsf{list}, \textsf{list},  \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{function}) $\rightarrow$ \textsf{function}\\
\textbf{fpminimax}(\emph{f}, \emph{monomials}, \emph{formats}, \emph{L}, \emph{indic1}, \emph{indic2}, \emph{indic3}, \emph{P}) : (\textsf{function}, \textsf{list}, \textsf{list}, \textsf{list},  \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{absolute$|$relative} $|$ \textsf{fixed$|$floating} $|$ \textsf{function}, \textsf{function}) $\rightarrow$ \textsf{function}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{f} is the function to be approximated
\item \emph{n} is the degree of the polynomial that must approximate \emph{f}
\item \emph{monomials} is the list of monomials that must be used to represent the polynomial that approximates~\emph{f}
\item \emph{formats} is a list indicating the formats that the coefficients of the polynomial must have
\item \emph{range} is the interval where the function must be approximated
\item \emph{L} is a list of interpolation points used by the method
\item \emph{indic1} (optional) is one of the optional indication parameters. See the detailed description below.
\item \emph{indic2} (optional) is one of the optional indication parameters. See the detailed description below.
\item \emph{indic3} (optional) is one of the optional indication parameters. See the detailed description below.
\item \emph{P} (optional) is the minimax polynomial to be considered for solving the problem.
\end{itemize}
\noindent Description: \begin{itemize}

\item \textbf{fpminimax} uses a heuristic (but practically efficient) method to find a
   good polynomial approximation of a function \emph{f} on an interval \emph{range}. It
   implements the method published in the article:\\
   Efficient polynomial $L^\infty$-approximations\\ 
   Nicolas Brisebarre and Sylvain Chevillard\\
   Proceedings of the 18th IEEE Symposium on Computer Arithmetic (ARITH 18)\\
   pp. 169-176

\item The basic usage of this command is \textbf{fpminimax}(\emph{f}, \emph{n}, \emph{formats}, \emph{range}).
   It computes a polynomial approximation of $f$ with degree at most $n$
   on the interval \emph{range}. \emph{formats} is a list of integers or format types 
   (such as \textbf{double}, \textbf{doubledouble}, etc.). The polynomial returned by the
   command has its coefficients that fit the formats indications. For 
   instance, if formats[0] is 35, the coefficient of degree 0 of the 
   polynomial will fit a floating-point format of 35 bits. If formats[1] 
   is D, the coefficient of degree 1 will be representable by a floating-point
   number with a precision of 53 bits (which is not necessarily an IEEE 754
   double precision number. See the remark below), etc.

\item The second argument may be either an integer or a list of integers
   interpreted as the list of desired monomials. For instance, the list
   $[|0,\,2,\,4,\,6|]$ indicates that the polynomial must be even and of
   degree at most 6. Giving an integer $n$ as second argument is equivalent
   as giving $[|0,\,\dots,\,n|]$.\\
   The list of formats is interpreted with respect to the list of monomials. For
   instance, if the list of monomials is $[|0,\,2,\,4,\,6|]$ and the list
   of formats is $[|161,\,107,\,53,\,24|]$, the coefficients of degree 0 is 
   searched as a floating-point number with precision 161, the coefficient of 
   degree 2 is searched as a number of precision 107, and so on.

\item The list of formats may contain either integers or format types
   (\textbf{halfprecision}, \textbf{single}, \textbf{double}, \textbf{doubledouble}, \textbf{tripledouble}, \textbf{doubleextended}
   and \textbf{quad}). The list may be too large or even infinite. Only the first
   indications will be considered. For instance, for a degree $n$ polynomial,
   $\mathrm{formats}[n+1]$ and above will be discarded. This lets one use elliptical
   indications for the last coefficients.

\item The floating-point coefficients considered by \textbf{fpminimax} do not have an
   exponent range. In particular, in the format list, \textbf{double} is an exact synonym
   for 53. Currently, \textbf{fpminimax} only ensures that the corresponding coefficient
   has at most 53 bits of mantissa. It does not imply that it is an
   IEEE-754 double.

\item By default, the list of formats is interpreted as a list of floating-point
   formats. This may be changed by passing \textbf{fixed} as an optional argument (see
   below). Let us take an example: \textbf{fpminimax}(f, 2, [107, DD, 53], [0;1]).
   Here the optional argument is missing (we could have set it to \textbf{floating}).
   Thus, \textbf{fpminimax} will search for a polynomial of degree 2 with a constant 
   coefficient that is a 107 bits floating-point number, etc.\\
   Currently, \textbf{doubledouble} is just a synonym for 107 and \textbf{tripledouble} a 
   synonym for 161. This behavior may change in the future (taking into
   account the fact that some double-doubles are not representable with
   107 bits).\\
   Second example: \textbf{fpminimax}(f, 2, [25, 18, 30], [0;1], \textbf{fixed}).
   In this case, \textbf{fpminimax} will search for a polynomial of degree 2 with a
   constant coefficient of the form $m/2^{25}$ where $m$ is an
   integer. In other words, it is a fixed-point number with 25 bits after
   the point. Note that even with argument \textbf{fixed}, the formats list is 
   allowed to contain \textbf{halfprecision}, \textbf{single}, \textbf{double}, \textbf{doubleextended}, 
   \textbf{doubledouble}, \textbf{quad} or \textbf{tripledouble}. In this this case, it is just 
   a synonym for 11, 24, 53, 64, 107, 113 or 161. This is deprecated and 
   may change in the future.

\item The fourth argument may be a range or a list. Lists are for advanced users
   that know what they are doing. The core of the  method is a kind of
   approximated interpolation. The list given here is a list of points that
   must be considered for the interpolation. It must contain at least as 
   many points as unknown coefficients. If you give a list, it is also 
   recommended that you provide the minimax polynomial as last argument.
   If you give a range, the list of points will be automatically computed.

\item The fifth, sixth and seventh arguments are optional. By default, \textbf{fpminimax}
   will approximate $f$ while optimizing the relative error, and interpreting
   the list of formats as a list of floating-point formats.\\
   This default behavior may be changed with these optional arguments. You
   may provide zero, one, two or three of the arguments in any order.
   This lets the user indicate only the non-default arguments.\\
   The three possible arguments are: \begin{itemize}
   \item \textbf{relative} or \textbf{absolute}: the error to be optimized;
   \item \textbf{floating} or \textbf{fixed}: formats of the coefficients;
   \item a constrained part $q$.
   \end{itemize}
   The constrained part lets the user assign in advance some of the
   coefficients. For instance, for approximating $\exp(x)$, it may
   be interesting to search for a polynomial $p$ of the form
                   $$p = 1 + x + \frac{x^2}{2} + a_3 x^3 + a_4 x^4.$$
   Thus, there is a constrained part $q = 1 + x + x^2/2$ and the unknown
   polynomial should be considered in the monomial basis $[|3, 4|]$.
   Calling \textbf{fpminimax} with monomial basis $[|3,\,4|]$ and constrained
   part $q$, will return a polynomial with the right form.

\item The last argument is for advanced users. It is the minimax polynomial that
   approximates the function $f$ in the monomial basis. If it is not given
   this polynomial will be automatically computed by \textbf{fpminimax}.
   \\
   This minimax polynomial is used to compute the list of interpolation
   points required by the method. In general, you do not have to provide this
   argument. But if you want to obtain several polynomials of the same degree
   that approximate the same function on the same range, just changing the
   formats, you should probably consider computing only once the minimax
   polynomial and the list of points instead of letting \textbf{fpminimax} recompute
   them each time.
   \\
   Note that in the case when a constrained part is given, the minimax 
   polynomial must take that into account. For instance, in the previous
   example, the minimax would be obtained by the following command:
          \begin{center}\verb~P = remez(1-(1+x+x^2/2)/exp(x), [|3,4|], range, 1/exp(x));~\end{center}
   Note that the constrained part is not to be added to $P$.

\item Note that \textbf{fpminimax} internally computes a minimax polynomial (using
   the same algorithm as \textbf{remez} command). Thus \textbf{fpminimax} may encounter
   the same problems as \textbf{remez}. In particular, it may be very slow 
   when Haar condition is not fulfilled. Another consequence is that
   currently \textbf{fpminimax} has to be run with a sufficiently high working precision.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> P = fpminimax(cos(x),6,[|DD, DD, D...|],[-1b-5;1b-5]);
> printexpansion(P);
(0x3ff0000000000000 + 0xbc09fda20235c100) + x * ((0x3b29ecd485d34781 + 0xb7c1cbc
97120359a) + x * (0xbfdfffffffffff98 + x * (0xbbfa6e0b3183cb0d + x * (0x3fa55555
55145337 + x * (0x3ca3540480618939 + x * 0xbf56c138142d8c3b)))))
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> P = fpminimax(sin(x),6,[|32...|],[-1b-5;1b-5], fixed, absolute);
> display = powers!;
> P;
x * (1 + x^2 * (-357913941 * 2^(-31) + x^2 * (35789873 * 2^(-32))))
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> P = fpminimax(exp(x), [|3,4|], [|D,24|], [-1/256; 1/246], 1+x+x^2/2);
> display = powers!;
> P;
1 + x * (1 + x * (1 * 2^(-1) + x * (375300225001191 * 2^(-51) + x * (5592621 * 2
^(-27)))))
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 4: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> f = cos(exp(x));
> pstar = remez(f, 5, [-1b-7;1b-7]);
> listpoints = dirtyfindzeros(f-pstar, [-1b-7; 1b-7]);
> P1 = fpminimax(f, 5, [|DD...|], listpoints, absolute, default, default, pstar)
;
> P2 = fpminimax(f, 5, [|D...|], listpoints, absolute, default, default, pstar);

> P3 = fpminimax(f, 5, [|D, D, D, 24...|], listpoints, absolute, default, defaul
t, pstar);
> print("Error of pstar: ", dirtyinfnorm(f-pstar, [-1b-7; 1b-7]));
Error of pstar:  7.9048441305459735102879831325718745399379329453102e-16
> print("Error of P1:    ", dirtyinfnorm(f-P1, [-1b-7; 1b-7]));
Error of P1:     7.9048441305459735159848647089192667442047469404883e-16
> print("Error of P2:    ", dirtyinfnorm(f-P2, [-1b-7; 1b-7]));
Error of P2:     8.2477144579950871061147021597406077993657714575238e-16
> print("Error of P3:    ", dirtyinfnorm(f-P3, [-1b-7; 1b-7]));
Error of P3:     1.08454277156993282593701156841863009789063333951055e-15
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{remez} (\ref{labremez}), \textbf{dirtyfindzeros} (\ref{labdirtyfindzeros}), \textbf{absolute} (\ref{lababsolute}), \textbf{relative} (\ref{labrelative}), \textbf{fixed} (\ref{labfixed}), \textbf{floating} (\ref{labfloating}), \textbf{default} (\ref{labdefault}), \textbf{halfprecision} (\ref{labhalfprecision}), \textbf{single} (\ref{labsingle}), \textbf{double} (\ref{labdouble}), \textbf{doubleextended} (\ref{labdoubleextended}), \textbf{doubledouble} (\ref{labdoubledouble}), \textbf{quad} (\ref{labquad}), \textbf{tripledouble} (\ref{labtripledouble}), \textbf{implementpoly} (\ref{labimplementpoly}), \textbf{coeff} (\ref{labcoeff}), \textbf{degree} (\ref{labdegree}), \textbf{roundcoefficients} (\ref{labroundcoefficients}), \textbf{guessdegree} (\ref{labguessdegree})
