\subsection{library}
\label{lablibrary}
\noindent Name: \textbf{library}\\
binds an external mathematical function to a variable in \sollya\\
\noindent Usage: 
\begin{center}
\textbf{library}(\emph{path}) : \textsf{string} $\rightarrow$ \textsf{function}\\
\end{center}
\noindent Description: \begin{itemize}

\item The command \textbf{library} lets you extend the set of mathematical
   functions known to \sollya.
   By default, \sollya knows the most common mathematical functions such
   as \textbf{exp}, \textbf{sin}, \textbf{erf}, etc. Within \sollya, these functions may be
   composed. This way, \sollya should satisfy the needs of a lot of
   users. However, for particular applications, one may want to
   manipulate other functions such as Bessel functions, or functions
   defined by an integral or even a particular solution of an ODE.

\item \textbf{library} makes it possible to let \sollya know about new functions. In
   order to let it know, you have to provide an implementation of the
   function you are interested in. This implementation is a C file containing
   a function of the form:
   \begin{verbatim} int my_ident(sollya_mpfi_t result, sollya_mpfi_t op, int n)\end{verbatim}
   The semantic of this function is the following: it is an implementation of
   the function and its derivatives in interval arithmetic.
   \verb|my_ident(result, I, n)| shall store in \verb|result| an enclosure 
   of the image set of the $n$-th derivative
   of the function f over \verb|I|: $f^{(n)}(I) \subseteq \mathrm{result}$.

\item The integer value returned by the function implementation currently has no meaning.

\item You do not need to provide a working implementation for any \verb|n|. Most functions
   of \sollya requires a relevant implementation only for $f$, $f'$ and $f''$. For higher 
   derivatives, its is not so critical and the implementation may just store 
   $[-\infty,\,+\infty]$ in result whenever $n>2$.

\item Note that you should respect somehow interval-arithmetic standards in your implementation:
   \verb|result| has its own precision and you should perform the 
   intermediate computations so that \verb|result| is as tight as possible.

\item You can include sollya.h in your implementation and use library 
   functionnalities of \sollya for your implementation. However, this requires to have compiled
   \sollya with \texttt{-fPIC} in order to make the \sollya executable code position 
   independent and to use a system on with programs, using \texttt{dlopen} to open
   dynamic routines can dynamically open themselves.

\item To bind your function into \sollya, you must use the same identifier as the
   function name used in your implementation file (\verb|my_ident| in the previous
   example). Once the function code has been bound to an identifier, you can use a simple assignment
   to assign the bound identifier to yet another identifier. This way, you may use convenient
   names inside \sollya even if your implementation environment requires you to use a less
   convenient name.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> bashexecute("gcc -fPIC -Wall -c libraryexample.c -I$HOME/.local/include");
> bashexecute("gcc -shared -o libraryexample libraryexample.o -lgmp -lmpfr");
> myownlog = library("./libraryexample");
> evaluate(log(x), 2);
0.69314718055994530941723212145817656807550013436024
> evaluate(myownlog(x), 2);
0.69314718055994530941723212145817656807550013436024
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{function} (\ref{labfunction}), \textbf{bashexecute} (\ref{labbashexecute}), \textbf{externalproc} (\ref{labexternalproc}), \textbf{externalplot} (\ref{labexternalplot}), \textbf{diff} (\ref{labdiff}), \textbf{evaluate} (\ref{labevaluate}), \textbf{libraryconstant} (\ref{lablibraryconstant})
