\subsection{max}
\label{labmax}
\noindent Name: \textbf{max}\\
determines which of given constant expressions has maximum value\\
\noindent Usage: 
\begin{center}
\textbf{max}(\emph{expr1},\emph{expr2},...,\emph{exprn}) : (\textsf{constant}, \textsf{constant}, ..., \textsf{constant}) $\rightarrow$ \textsf{constant}\\
\textbf{max}(\emph{l}) : \textsf{list} $\rightarrow$ \textsf{constant}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{expr} are constant expressions.
\item \emph{l} is a list of constant expressions.
\end{itemize}
\noindent Description: \begin{itemize}

\item \textbf{max} determines which of a given set of constant expressions
   \emph{expr} has maximum value. To do so, \textbf{max} tries to increase the
   precision used for evaluation until it can decide the ordering or some
   maximum precision is reached. In the latter case, a warning is printed
   indicating that there might actually be another expression that has a
   greater value.

\item Even though \textbf{max} determines the maximum expression by evaluation, it 
   returns the expression that is maximum as is, i.e. as an expression
   tree that might be evaluated to any accuracy afterwards.

\item \textbf{max} can be given either an arbitrary number of constant
   expressions in argument or a list of constant expressions. The list
   however must not be end-elliptic.

\item Users should be aware that the behavior of \textbf{max} follows the IEEE
   754-2008 standard with respect to NaNs. In particular, a NaN given as
   the first argument will not be promoted as a result unless the other
   argument is a NaN. This means that NaNs may seem to disappear during
   computations.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> max(1,2,3,exp(5),log(0.25));
1.48413159102576603421115580040552279623487667593878e2
> max(17);
17
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> l = [|1,2,3,exp(5),log(0.25)|];
> max(l);
1.48413159102576603421115580040552279623487667593878e2
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> print(max(exp(17),sin(62)));
exp(17)
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 4: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 1!;
> print(max(17 + log2(13)/log2(9),17 + log(13)/log(9)));
Warning: maximum computation relies on floating-point result that is faithfully 
evaluated and different faithful roundings toggle the result.
17 + log2(13) / log2(9)
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{min} (\ref{labmin}), \textbf{$==$} (\ref{labequal}), \textbf{!$=$} (\ref{labneq}), \textbf{$>=$} (\ref{labge}), \textbf{$>$} (\ref{labgt}), \textbf{$<$} (\ref{lablt}), \textbf{$<=$} (\ref{lable}), \textbf{in} (\ref{labin}), \textbf{inf} (\ref{labinf}), \textbf{sup} (\ref{labsup})
