\subsection{numberroots}
\label{labnumberroots}
\noindent Name: \textbf{numberroots}\\
Computes the number of roots of a polynomial in a given range.\\
\noindent Usage: 
\begin{center}
\textbf{numberroots}(\emph{p}, \emph{I}) : (\textsf{function}, \textsf{range}) $\rightarrow$ \textsf{integer}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{p} is a polynomial.
\item \emph{I} is an interval.
\end{itemize}
\noindent Description: \begin{itemize}

\item \textbf{numberroots} rigorously computes the number of roots of polynomial the $p$ in
   the interval $I$. The technique used is Sturm's algorithm. The value returned
   is not just a numerical estimation of the number of roots of $p$ in $I$: it is
   the exact number of roots.

\item The command \textbf{findzeros} computes safe enclosures of all the zeros of a
   function, without forgetting any, but it is not guaranteed to separate them
   all in distinct intervals. \textbf{numberroots} is more accurate since it guarantees 
   the exact number of roots. However, it does not compute them. It may be used,
   for instance, to certify that \textbf{findzeros} did not put two distinct roots in 
   the same interval.

\item Multiple roots are counted only once.

\item The interval $I$ must be bounded. The algorithm cannot handle unbounded
   intervals. Moreover, the interval is considered as a closed interval: if one
   (or both) of the endpoints of $I$ are roots of $p$, they are counted.

\item The argument $p$ can be any expression, but if \sollya fails to prove that
   it is a polynomial an error is produced. Also, please note that if the
   coefficients of $p$ or the endpoints of $I$ are not exactly representable,
   they are first numerically evaluated, before the algorithm is used. In that
   case, the counted number of roots corresponds to the rounded polynomial on
   the rounded interval \textbf{and not} to the exact parameters given by the user.
   A warning is displayed to inform the user.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> numberroots(1+x-x^2, [1,2]);
1
> findzeros(1+x-x^2, [1,2]);
[|[1.617919921875;1.6180419921875]|]
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> numberroots((1+x)*(1-x), [-1,1]);
2
> numberroots(x^2, [-1,1]);
1
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 1!;
> numberroots(x-pi, [0,4]);
Warning: the 0th coefficient of the polynomial is neither a floating point
constant nor can be evaluated without rounding to a floating point constant.
Will faithfully evaluate it with the current precision (165 bits) 
1
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 4: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 1!;
> numberroots(1+x-x^2, [0, @Inf@]);
Warning: the given interval must have finite bounds.
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
error
> numberroots(exp(x), [0, 1]);
Warning: the given function must be a polynomial in this context.
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
error
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{dirtyfindzeros} (\ref{labdirtyfindzeros}), \textbf{findzeros} (\ref{labfindzeros})
