\subsection{plot}
\label{labplot}
\noindent Name: \textbf{plot}\\
plots one or several functions\\
\noindent Usage: 
\begin{center}
\textbf{plot}(\emph{f1}, ... ,\emph{fn}, \emph{I}) : (\textsf{function}, ... ,\textsf{function}, \textsf{range}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{f1}, ... ,\emph{fn}, \emph{I}, \textbf{file}, \emph{name}) : (\textsf{function}, ... ,\textsf{function}, \textsf{range}, \textbf{file}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{f1}, ... ,\emph{fn}, \emph{I}, \textbf{postscript}, \emph{name}) : (\textsf{function}, ... ,\textsf{function}, \textsf{range}, \textbf{postscript}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{f1}, ... ,\emph{fn}, \emph{I}, \textbf{postscriptfile}, \emph{name}) : (\textsf{function}, ... ,\textsf{function}, \textsf{range}, \textbf{postscriptfile}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{L}, \emph{I}) : (\textsf{list}, \textsf{range}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{L}, \emph{I}, \textbf{file}, \emph{name}) : (\textsf{list}, \textsf{range}, \textbf{file}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{L}, \emph{I}, \textbf{postscript}, \emph{name}) : (\textsf{list}, \textsf{range}, \textbf{postscript}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\textbf{plot}(\emph{L}, \emph{I}, \textbf{postscriptfile}, \emph{name}) : (\textsf{list}, \textsf{range}, \textbf{postscriptfile}, \textsf{string}) $\rightarrow$ \textsf{void}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{f1}, ..., \emph{fn} are functions to be plotted.
\item \emph{L} is a list of functions to be plotted.
\item \emph{I} is the interval where the functions have to be plotted.
\item \emph{name} is a string representing the name of a file.
\end{itemize}
\noindent Description: \begin{itemize}

\item This command plots one or several functions \emph{f1}, ... ,\emph{fn} on an interval \emph{I}.
   Functions can be either given as parameters of \textbf{plot} or as a list \emph{L}
   which elements are functions.
   The functions are drawn on the same plot with different colors.

\item If \emph{L} contains an element that is not a function (or a constant), an error
   occurs.

\item \textbf{plot} relies on the value of global variable \textbf{points}. Let $n$ be the 
   value of this variable. The algorithm is the following: each function is 
   evaluated at $n$ evenly distributed points in \emph{I}. At each point, the 
   computed value is a faithful rounding of the exact value with a sufficiently
   high precision. Each point is finally plotted.
   This should avoid numerical artefacts such as critical cancellations.

\item You can save the function plot either as a data file or as a postscript file.

\item If you use argument \textbf{file} with a string \emph{name}, \sollya will save a data file
   called name.dat and a gnuplot directives file called name.p. Invoking gnuplot
   on name.p will plot the data stored in name.dat.

\item If you use argument \textbf{postscript} with a string \emph{name}, \sollya will save a 
   postscript file called name.eps representing your plot.

\item If you use argument \textbf{postscriptfile} with a string \emph{name}, \sollya will 
   produce the corresponding name.dat, name.p and name.eps.

\item This command uses gnuplot to produce the final plot.
   If your terminal is not graphic (typically if you use \sollya through 
   ssh without -X)
   gnuplot should be able to detect that and produce an ASCII-art version on the
   standard output. If it is not the case, you can either store the plot in a
   postscript file to view it locally, or use \textbf{asciiplot} command.

\item If every function is constant, \textbf{plot} will not plot them but just display
   their value.

\item If the interval is reduced to a single point, \textbf{plot} will just display the
   value of the functions at this point.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> plot(sin(x),0,cos(x),[-Pi,Pi]);
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> plot(sin(x),0,cos(x),[-Pi,Pi],postscriptfile,"plotSinCos");
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> plot(exp(0), sin(1), [0;1]);
1
0.84147098480789650665250232163029899962256306079837
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 4: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> plot(sin(x), cos(x), [1;1]);
0.84147098480789650665250232163029899962256306079837
0.54030230586813971740093660744297660373231042061792
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{externalplot} (\ref{labexternalplot}), \textbf{asciiplot} (\ref{labasciiplot}), \textbf{file} (\ref{labfile}), \textbf{postscript} (\ref{labpostscript}), \textbf{postscriptfile} (\ref{labpostscriptfile}), \textbf{points} (\ref{labpoints})
