\subsection{proc}
\label{labproc}
\noindent Name: \textbf{proc}\\
defines a \sollya procedure\\
\noindent Usage: 
\begin{center}
\textbf{proc}(\emph{formal parameter1}, \emph{formal parameter2},..., \emph{formal parameter n}) \key{$\lbrace$} \emph{procedure body} \key{$\rbrace$} : \textsf{void} $\rightarrow$ \textsf{procedure}\\
\textbf{proc}(\emph{formal parameter1}, \emph{formal parameter2},..., \emph{formal parameter n}) \key{$\lbrace$} \emph{procedure body} \textbf{return} \emph{expression}; \key{$\rbrace$} : \textsf{void} $\rightarrow$ \textsf{procedure}\\
\textbf{proc}(\emph{formal list parameter} = ...) \key{$\lbrace$} \emph{procedure body} \key{$\rbrace$} : \textsf{void} $\rightarrow$ \textsf{procedure}\\
\textbf{proc}(\emph{formal list parameter} = ...) \key{$\lbrace$} \emph{procedure body} \textbf{return} \emph{expression}; \key{$\rbrace$} : \textsf{void} $\rightarrow$ \textsf{procedure}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{formal parameter1}, \emph{formal parameter2} through \emph{formal parameter n} represent identifiers used as formal parameters
\item \emph{formal list parameter} represents an identifier used as a formal parameter for the list of an arbitrary number of parameters
\item \emph{procedure body} represents the imperative statements in the body of the procedure
\item \emph{expression} represents the expression \textbf{proc} shall evaluate to
\end{itemize}
\noindent Description: \begin{itemize}

\item The \textbf{proc} keyword allows for defining procedures in the \sollya
   language. These procedures are common \sollya objects that can be
   applied to actual parameters after definition. Upon such an
   application, the \sollya interpreter applies the actual parameters to
   the formal parameters \emph{formal parameter1} through \emph{formal parameter n}
   (resp. builds up the list of arguments and applies it to the list
   \emph{formal list parameter}) and executes the \emph{procedure body}. The
   procedure applied to actual parameters evaluates then to the
   expression \emph{expression} in the \textbf{return} statement after the \emph{procedure body} 
   or to \textbf{void}, if no return statement is given (i.e. a \textbf{return}
   \textbf{void} statement is implicitly given).

\item \sollya procedures defined by \textbf{proc} have no name. They can be bound
   to an identifier by assigning the procedure object a \textbf{proc}
   expression produces to an identifier. However, it is possible to use
   procedures without giving them any name. For instance, \sollya
   procedures, i.e. procedure objects, can be elements of lists. They can
   even be given as an argument to other internal \sollya procedures. See
   also \textbf{procedure} on this subject.

\item Upon definition of a \sollya procedure using \textbf{proc}, no type check
   is performed. More precisely, the statements in \emph{procedure body} are
   merely parsed but not interpreted upon procedure definition with
   \textbf{proc}. Type checks are performed once the procedure is applied to
   actual parameters or to \textbf{void}. At this time, if the procedure was
   defined using several different formal parameters \emph{formal parameter 1}
   through \emph{formal parameter n}, it is checked whether the number of
   actual parameters corresponds to the number of formal parameters. If
   the procedure was defined using the syntax for a procedure with an
   arbitrary number of parameters by giving a \emph{formal list parameter},
   the number of actual arguments is not checked but only a list <formal
   list parameter> of appropriate length is built up. Type checks are
   further performed upon execution of each statement in \emph{procedure body}
   and upon evaluation of the expression \emph{expression} to be returned.
    
   Procedures defined by \textbf{proc} containing a \textbf{quit} or \textbf{restart} command
   cannot be executed (i.e. applied). Upon application of a procedure,
   the \sollya interpreter checks beforehand for such a statement. If one
   is found, the application of the procedure to its arguments evaluates
   to \textbf{error}. A warning is displayed. Remark that in contrast to other
   type or semantic correctness checks, this check is really performed
   before interpreting any other statement in the body of the procedure.

\item Through the \textbf{var} keyword it is possible to declare local
   variables and thus to have full support of recursive procedures. This
   means a procedure defined using \textbf{proc} may contain in its \emph{procedure body} 
   an application of itself to some actual parameters: it suffices
   to assign the procedure (object) to an identifier with an appropriate
   name.

\item \sollya procedures defined using \textbf{proc} may return other
   procedures. Further \emph{procedure body} may contain assignments of
   locally defined procedure objects to identifiers. See \textbf{var} for the
   particular behaviour of local and global variables.

\item The expression \emph{expression} returned by a procedure is evaluated with
   regard to \sollya commands, procedures and external
   procedures. Simplification may be performed.  However, an application
   of a procedure defined by \textbf{proc} to actual parameters evaluates to the
   expression \emph{expression} that may contain the free global variable or
   that may be composed.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> succ = proc(n) { return n + 1; };
> succ(5);
6
> 3 + succ(0);
4
> succ;
proc(n)
{
nop;
return (n) + (1);
}
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 2: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> add = proc(m,n) { var res; res := m + n; return res; };
> add(5,6);
11
> add;
proc(m, n)
{
var res;
res := (m) + (n);
return res;
}
> verbosity = 1!;
> add(3);
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
error
> add(true,false);
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
Warning: the given expression or command could not be handled.
error
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 3: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> succ = proc(n) { return n + 1; };
> succ(5);
6
> succ(x);
1 + x
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 4: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> hey = proc() { print("Hello world."); };
> hey();
Hello world.
> print(hey());
Hello world.
void
> hey;
proc()
{
print("Hello world.");
return void;
}
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 5: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> fac = proc(n) { var res; if (n == 0) then res := 1 else res := n * fac(n - 1);
 return res; };
> fac(5);
120
> fac(11);
39916800
> fac;
proc(n)
{
var res;
if (n) == (0) then
res := 1
else
res := (n) * (fac((n) - (1)));
return res;
}
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 6: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> myprocs = [| proc(m,n) { return m + n; }, proc(m,n) { return m - n; } |];
> (myprocs[0])(5,6);
11
> (myprocs[1])(5,6);
-1
> succ = proc(n) { return n + 1; };
> pred = proc(n) { return n - 1; };
> applier = proc(p,n) { return p(n); };
> applier(succ,5);
6
> applier(pred,5);
4
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 7: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity = 1!;
> myquit = proc(n) { print(n); quit; };
> myquit;
proc(n)
{
print(n);
quit;
return void;
}
> myquit(5);
Warning: a quit or restart command may not be part of a procedure body.
The procedure will not be executed.
Warning: an error occurred while executing a procedure.
Warning: the given expression or command could not be handled.
error
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 8: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> printsucc = proc(n) { var succ; succ = proc(n) { return n + 1; }; print("Succe
ssor of",n,"is",succ(n)); };
> printsucc(5);
Successor of 5 is 6
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 9: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> makeadd = proc(n) { var add; print("n =",n); add = proc(m,n) { return n + m; }
; return add; };
> makeadd(4);
n = 4
proc(m, n)
{
nop;
return (n) + (m);
}
> (makeadd(4))(5,6);
n = 4
11
\end{Verbatim}
\end{minipage}\end{center}
\noindent Example 10: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> sumall = proc(L = ...) { var acc, i; acc = 0; for i in L do acc = acc + i; ret
urn acc; };
> sumall;
proc(L = ...)
{
var acc, i;
acc = 0;
for i in L do
acc = (acc) + (i);
return acc;
}
> sumall();
0
> sumall(2);
2
> sumall(2,5);
7
> sumall(2,5,7,9,16);
39
> sumall @ [|1,...,10|];
55
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{return} (\ref{labreturn}), \textbf{externalproc} (\ref{labexternalproc}), \textbf{void} (\ref{labvoid}), \textbf{quit} (\ref{labquit}), \textbf{restart} (\ref{labrestart}), \textbf{var} (\ref{labvar}), \textbf{@} (\ref{labconcat}), \textbf{error} (\ref{laberror})
