\subsection{roundingwarnings}
\label{labroundingwarnings}
\noindent Name: \textbf{roundingwarnings}\\
global variable controlling whether or not a warning is displayed when roundings occur.\\
\noindent Usage: 
\begin{center}
\textbf{roundingwarnings} = \emph{activation value} : \textsf{on$|$off} $\rightarrow$ \textsf{void}\\
\textbf{roundingwarnings} = \emph{activation value} ! : \textsf{on$|$off} $\rightarrow$ \textsf{void}\\
\textbf{roundingwarnings} : \textsf{on$|$off}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{activation value} controls if warnings should be shown or not
\end{itemize}
\noindent Description: \begin{itemize}

\item \textbf{roundingwarnings} is a global variable. When its value is \textbf{on}, warnings are
   emitted in appropriate verbosity modes (see \textbf{verbosity}) when roundings
   occur.  When its value is \textbf{off}, these warnings are suppressed.

\item This mode depends on a verbosity of at least 1. See
   \textbf{verbosity} for more details.

\item Default is \textbf{on} when the standard input is a terminal and
   \textbf{off} when \sollya input is read from a file.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity=1!;
> roundingwarnings = on;
Rounding warning mode has been activated.
> exp(0.1);
Warning: Rounding occurred when converting the constant "0.1" to floating-point 
with 165 bits.
If safe computation is needed, try to increase the precision.
Warning: rounding has happened. The value displayed is a faithful rounding of th
e true result.
1.1051709180756476248117078264902466682245471947375
> roundingwarnings = off;
Rounding warning mode has been deactivated.
> exp(0.1);
1.1051709180756476248117078264902466682245471947375
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{on} (\ref{labon}), \textbf{off} (\ref{laboff}), \textbf{verbosity} (\ref{labverbosity}), \textbf{midpointmode} (\ref{labmidpointmode}), \textbf{rationalmode} (\ref{labrationalmode})
