\subsection{verbosity}
\label{labverbosity}
\noindent Name: \textbf{verbosity}\\
global variable controlling the amount of information displayed by commands.\\
\noindent Usage: 
\begin{center}
\textbf{verbosity} = \emph{n} : \textsf{integer} $\rightarrow$ \textsf{void}\\
\textbf{verbosity} = \emph{n} ! : \textsf{integer} $\rightarrow$ \textsf{void}\\
\textbf{verbosity} : \textsf{integer}\\
\end{center}
Parameters: 
\begin{itemize}
\item \emph{n} controls the amount of information to be displayed
\end{itemize}
\noindent Description: \begin{itemize}

\item \textbf{verbosity} accepts any integer value. At level 0, commands do not display anything
   on standard output. Note that very critical information may however be displayed on
   standard error.

\item Default level is 1. It displays important information such as warnings when 
   roundings happen.

\item For higher levels more information is displayed depending on the command.
\end{itemize}
\noindent Example 1: 
\begin{center}\begin{minipage}{15cm}\begin{Verbatim}[frame=single]
> verbosity=0!;
> 1.2+"toto";
error
> verbosity=1!;
> 1.2+"toto";
Warning: Rounding occurred when converting the constant "1.2" to floating-point 
with 165 bits.
If safe computation is needed, try to increase the precision.
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
error
> verbosity=2!;
> 1.2+"toto";
Warning: Rounding occurred when converting the constant "1.2" to floating-point 
with 165 bits.
If safe computation is needed, try to increase the precision.
Warning: at least one of the given expressions or a subexpression is not correct
ly typed
or its evaluation has failed because of some error on a side-effect.
Information: the expression or a partial evaluation of it has been the following
:
(1.19999999999999999999999999999999999999999999999999) + ("toto")
error
\end{Verbatim}
\end{minipage}\end{center}
See also: \textbf{roundingwarnings} (\ref{labroundingwarnings})
